Constructing Statistical Deformable Models with Shape Flow

### 2015-04-15 contents/introduction.tex
Update Abstract & Evalutation Structure
Introduction edited

Uses Introductions from Nondas' paper "Automatic Construction of Deformable Models In-The-Wild" as place holder

Many deformable objects exist everywhere around us.
Some examples are the human face and body, animals, vehicles
such as cars and motorcycles and objects of everyday
use like tables, chairs etc. Most of these objects consist of
different parts and appear in instances with great variance
in shape and appearance. Thus, the concept of a deformable
object refers to the deformation of both shape and appearance
of an object. For example cars have parts (i.e. doors,
windows, tires etc.) with significant changes in shape, size
and texture. Furthermore, faces consist of parts (i.e. nose,
eyes, mouth etc.) which not only vary with respect to shape
and appearance, but can demonstrate a number of expressions
due to muscles. Recently, we have witnessed a great
progress in object detection, alignment and recognition.

In order to train deformable models with good generalization
ability, a large amount of carefully annotated data
is needed. Developing useful datasets and benchmarks that
can contribute in the progress of an application domain is
a highly time consuming and costly procedure. It requires
both careful selection of the images, so that they can model
the vast amount of an object’s variability, and careful annotation
of the various parts of the object (or landmarks).
The amount of annotation that is required depends on both
the object and the application. In faces, for example, where
many landmark points are needed in tasks such as facial expression
analysis, motion capture and expression transfer,
usually more than 60 points are annotated \cite{?}.
To illustrate how much time consuming careful face annotation
is, according to our experience, a trained annotator
may need an average of 8 minutes per image for the manual
annotation of 68 landmarks1. This means that the annotation
of 1000 images requires a total of about 130 hours2.
Furthermore, fatigue can cause errors on the accuracy and
consistency of annotations and they may require correction.

>>>>

Building and training deformable models of objects (e.g. faces, ears, bodies, bottles, vehicles etc.) has recently been widely researched for object detection, part localisation, fitting, recognition and tracking using manual annotated data. Shapes and appearances of majority of objects vary significantly from object instance. To illustrate, ears have complicated inner structures (e.g. helix, crus antihelicis, scapha, tragus, lobe etc.) which differs remarkably between ears and highly sensitive to intensity changes. Recently, we have witnessed a great
progress in object detection, alignment and recognition.

To train deformable models with satisfiable generalisation, large amount of careful manual annotation is required. However, landmark annotation is extremely time consuming, work intensive and, most importantly, number of landmarks have to stay consistent for all training data while it is challenging to maintain the consistency or to avoid bias when annotating objects that having reach features. In ears, for example, large amount of landmarks required as the inner structure of ears are complex. To annotate an ear with clear contour, helix and cavum conchae, more than 50 landmarks are required. The inconsistency of ear features enhanced the difficulty of accurate annotation (e.g. Fossa Triangularies and Crus Helicis are optional feature of an ear also their visibility is highly sensitive to pose and illumination variation).
To illustrate how much time consuming careful ear annotation is, according to our experience, a trained annotator may need an average of (TODO: experiment) minutes per image for the manual annotation of 55 landmarks. This means that the annotation of 1000 images requires a total of about (TODO: experiment). Furthermore, personal issue, prior knowledge and circumstances can heavily bias the annotation accuracy thus second round correction might desired.

We propose framework to handle training data with inconsistent annotation, build dense shape model and introduce effective curve annotation methods. In our experiment, we show building dense models trained with inconsistent annotation and trained with curve annotation improves performance of discriminative models trained on carefully annotated data.

Uses Introductions from Nondas' paper "Automatic Construction of Deformable Models In-The-Wild" as place holder.
